```{r setup, include=FALSE}
#| echo: false
#| warning: false

library(reticulate)
```

```{python}
#| echo: false
#| output: false
import sys
from pathlib import Path

sys.path.append(str(Path.cwd().resolve().parent))

from content.python.api_calls_v2 import get_habitat_stat_value, get_structure_count_value

connected_habitat_km = get_habitat_stat_value(
    "CHETI", "as", "connected_naturally_accessible_habitat_km"
)
disconnected_habitat_km = get_habitat_stat_value(
    "CHETI", "as", "disconnected_naturally_accessible_habitat_km"
)
total_habitat_km = get_habitat_stat_value("CHETI", "as", "total_habitat_km")
connectivity_status_pct = get_habitat_stat_value(
    "CHETI", "as", "connectivity_status", digits=1
)

potentially_disconnect_structures = get_structure_count_value(
    "CHETI", "as", "structures_that_potentially_disconnect_habitat"
)
further_field_assessment = get_structure_count_value(
    "CHETI", "as", "require_further_assessment"
)
```

# Connectivity Status Results {-}

## Current Connectivity Status {-}

In the NE PEI watersheds, 103.74 km of Atlantic Salmon habitat are currently connected to the ocean and 98.37 km are disconnected from it. This means that 51.34% of the 202.11 km of total salmon habitat is connected. 

In the NE PEI watersheds, 43.04 km of American Eel habitat are currently connected to the ocean and 132.77 km are disconnected from it. This means that 24.48% of the 176.80 km of total eel habitat is connected. 

In the NE PEI watersheds, X km of Rainbow Smelt habitat are currently connected to the ocean and X km are disconnected from it. This means that X% of the X km of total smelt habitat is connected. 



## Structure Count {-}

In the NE PEI watersheds, X structures potentially disconnect Atlantic Salmon and American Eel habitat. Of these, 10 are identified as barriers in need of rehabilitation (priority barriers), 0 are identified as barriers that do not warrant rehabilitation (non-actionable), and 4 require further field assessment. 

In the NE PEI watersheds, X structures potentially disconnect Rainbow Smelt. Of these, X are identified as barriers in need of rehabilitation (priority barriers), X are identified as barriers that do not warrant rehabilitation (non-actionable), and X require further field assessment. 


## Map {-}
!["Figure 1: Map of Atlantic Salmon habitat and structures that are confirmed or potential barriers to fish passage in the NE PEI watersheds as of July 2026. Structure data were obtained from PEIFishPass and the Canadian Aquatic Barriers Database (aquaticbarriers.ca). The accessibility model represents areas of the watershed that Atlantic Salmon could access naturally in the absence of anthropogenic barriers. The habitat model represents the subset of accessible waterbodies that may be used by Atlantic Salmon for spawning and rearing. Available local knowledge and data were incorporated and overruled habitat model results. Thick orange lines represent partially disconnected habitat (upstream of a partial barrier), and thick red lines represent habitat considered to be fully disconnected (upstream of barriers or unassessed structures). Barriers that were rehabilitated through implementation of this plan are shown, but other excluded structures (e.g., those found to be passable) are not. "](images/Figure1.png){fig-cap-location="top"}

!["Figure 2: Map of American Eel habitat and structures that are confirmed or potential barriers to fish passage in the NE PEI watersheds as of July 2026. Structure data were obtained from PEIFishPass and the Canadian Aquatic Barriers Database (aquaticbarriers.ca). The accessibility model represents areas of the watershed that American Eel could access naturally in the absence of anthropogenic barriers. The habitat model represents the subset of accessible waterbodies that may be used by American Eel for feeding and growth. Available local knowledge and data were incorporated and overruled habitat model results. Thick orange lines represent partially disconnected habitat (upstream of a partial barrier), and thick red lines represent habitat considered to be fully disconnected (upstream of barriers or unassessed structures). Barriers that were rehabilitated through implementation of this plan are shown, but other excluded structures (e.g., those found to be passable) are not. "](images/Figure2.png){fig-cap-location="top"}

Figure 3: Map of Rainbow Smelt habitat and structures that are confirmed or potential barriers to fish passage in the NE PEI watersheds as of July 2026. Structure data were obtained from PEIFishPass and the Canadian Aquatic Barriers Database (aquaticbarriers.ca). The accessibility model represents areas of the watershed that Rainbow Smelt could access naturally in the absence of anthropogenic barriers. The habitat model represents the subset of accessible waterbodies that may be used by Rainbow Smelt for spawning and rearing. Available local knowledge and data were incorporated and overruled habitat model results. Thick orange lines represent partially disconnected habitat (upstream of a partial barrier), and thick red lines represent habitat considered to be fully disconnected (upstream of barriers or unassessed structures). Barriers that were rehabilitated through implementation of this plan are shown, but other excluded structures (e.g., those found to be passable) are not. 


## Habitat Accumulation Curve {-}

!["Figure 4: Habitat Accumulation Curve (HAC) showing all structures (n=129) in the NE PEI watersheds that potentially block Atlantic Salmon habitat as of July 2026. Structures are ranked based on how much (km) weighted habitat for Atlantic Salmon is upstream. The y-axis represents the cumulative amount of potential habitat gain (km), down-weighted by 75% for first-order streams and 25% for second-order streams, and by passability (25%, 50%, or 75%) of partial barriers. Sets of structures are indicated by a grey line below the curve, indicating that gains from addressing the downstream structure alone would be less than the combined gains from addressing all structures in the set. Habitat upstream of unassessed structures is considered disconnected until field assessments are completed. Structures that were excluded as passable are not shown; however, rehabilitated barriers are shown to demonstrate the connectivity gains achieved through implementation of this plan.  "](images/Figure4.png){fig-cap-location="top"}

!["Figure 5: Habitat Accumulation Curve (HAC) showing top-ranked structures (n=30) in the NE PEI watersheds that potentially block Atlantic Salmon habitat as of July 2026. Structures are ranked based on how much (km) weighted habitat for Atlantic Salmon is upstream. The y-axis represents the cumulative amount of potential habitat gain (km), down-weighted by 75% for first-order streams and 25% for second-order streams, and by passability (25%, 50%, or 75%) of partial barriers. Sets of structures are indicated by a grey line below the curve, indicating that gains from addressing the downstream structure alone would be less than the combined gains from addressing all structures in the set. Habitat upstream of unassessed structures is considered disconnected until field assessments are completed. Structures that were excluded as passable are not shown; however, rehabilitated barriers are shown to demonstrate the connectivity gains achieved through implementation of this plan. "](images/Figure5.png){fig-cap-location="top"}

!["Figure 6: Habitat Accumulation Curve (HAC) showing all structures (n=77) in the NE PEI watersheds that potentially block American Eel habitat as of July 2026. Structures are ranked based on how much (km) weighted habitat for American Eel is upstream. The y-axis represents the cumulative amount of potential habitat gain (km), down-weighted by 75% for first-order streams and 25% for second-order streams, and by passability (25%, 50%, or 75%) of partial barriers. Sets of structures are indicated by a grey line below the curve, indicating that gains from addressing the downstream structure alone would be less than the combined gains from addressing all structures in the set. Habitat upstream of unassessed structures is considered disconnected until field assessments are completed. Structures that were excluded as passable are not shown; however, rehabilitated barriers are shown to demonstrate the connectivity gains achieved through implementation of this plan. "](images/Figure6.png){fig-cap-location="top"}

!["Figure 7: Habitat Accumulation Curve (HAC) showing the top-ranked structures (n=30) in the NE PEI watersheds that potentially block American Eel habitat as of July 2026. Structures are ranked based on how much (km) weighted habitat for American Eel is upstream. The y-axis represents the cumulative amount of potential habitat gain (km), down-weighted by 75% for first-order streams and 25% for second-order streams, and by passability (25%, 50%, or 75%) of partial barriers. Sets of structures are indicated by a grey line below the curve, indicating that gains from addressing the downstream structure alone would be less than the combined gains from addressing all structures in the set. Habitat upstream of unassessed structures is considered disconnected until field assessments are completed. Structures that were excluded as passable are not shown; however, rehabilitated barriers are shown to demonstrate the connectivity gains achieved through implementation of this plan. "](images/Figure7.png){fig-cap-location="top"}

Figure 8: Habitat Accumulation Curve (HAC) showing top-ranked structures (XX) in the NE PEI watersheds that potentially block Rainbow Smelt habitat as of July 2026. Structures are ranked based on how much (km) weighted habitat for Rainbow Smelt is upstream. The y-axis represents the cumulative amount of potential habitat gain (km), down-weighted by 75% for first-order streams and 25% for second-order streams, and by passability (25%, 50%, or 75%) of partial barriers. Sets of structures are indicated by a grey line below the curve, indicating that gains from addressing the downstream structure alone would be less than the combined gains from addressing all structures in the set. Habitat upstream of unassessed structures is considered disconnected until field assessments are completed. Structures that were excluded as passable are not shown; however, rehabilitated barriers are shown to demonstrate the connectivity gains achieved through implementation of this plan. 


## Habitat Accumulation Summary Table

```{r}
#| label: table-1
#| tbl-cap: "Table 1: Labelled structures in the Habitat Accumulation Curve (Figure 4), their rank, and weighted upstream habitat for Atlantic Salmon in NE PEI watersheds. Habitat gain (averaged across sets) represents the amount (km) of weighted habitat upstream of that structure to the next structure and depicts the potential gain in habitat if passage at that structure was restored. "
#| warning: false
#| echo: false
source("Rscripts/table_formatting.R")
library("flextable")

data <- read.csv("data/Table1.csv", check.names = FALSE)
ft <- flextable(data)
ft <- format_flextable(ft)
ft
```

```{r}
#| label: table-2
#| tbl-cap: "Table 2: Labelled structures in the Habitat Accumulation Curve (Figure 5), their rank, and weighted upstream habitat for American Eel in NE PEI watersheds. Habitat gain (averaged across sets) represents the amount (km) of weighted habitat upstream of that structure to the next structure and depicts the potential gain in habitat if passage at that structure was restored.  "
#| warning: false
#| echo: false
source("Rscripts/table_formatting.R")
library("flextable")

data <- read.csv("data/Table2.csv", check.names = FALSE)
ft <- flextable(data)
ft <- format_flextable(ft)
ft
```

Table 3: Labelled structures in the Habitat Accumulation Curve (Figure 6), their rank, and weighted upstream habitat for Rainbow Smelt in NE PEI watersheds. Habitat gain (averaged across sets) represents the amount (km) of weighted habitat upstream of that structure to the next structure and depicts the potential gain in habitat if passage at that structure was restored.  
```{r}
#| label: table-3
#| tbl-cap: "Table 3: Labelled structures in the Habitat Accumulation Curve (Figure 6), their rank, and weighted upstream habitat for Rainbow Smelt in NE PEI watersheds. Habitat gain (averaged across sets) represents the amount (km) of weighted habitat upstream of that structure to the next structure and depicts the potential gain in habitat if passage at that structure was restored.  "
#| warning: false
#| echo: false
#source("Rscripts/table_formatting.R")
#library("flextable")

#data <- read.csv("data/Table3.csv", check.names = FALSE)
#ft <- flextable(data)
#ft <- format_flextable(ft)
#ft
```

